An efficient acyclic contact planner for multiped robots

### 2017-01-16 body/intro.tex
updates

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the combinatorial of the contact selection problem. First we introduce a low-dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling presents drawbacks and advantages discussed in previous related literature, which we summarize in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected to the contact manifold using an inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. \citeauthor{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for a humanoid character, and managed to obtain dynamically-consistent motions
at \gls{interactive} frame rates (we say that a method is interactive when the computation time for planning is less than the
time for execution). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

>>>>

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the combinatorial of the contact selection problem. First we introduce a low-dimensional space, called the \textit{contact reachable} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact reachable} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling presents drawbacks and advantages discussed in previous related literature, which we summarize in the following.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, that is that for each root configuration in the guide path, there exists a joint configuration that results in static equilibrium. They did it by extending key frames of the path into whole-body configurations using a generalized inverse kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. \citeauthor{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
used black-box physics simulators to perform Model Predictive Control for an avatar, and obtained dynamically-consistent motions
at \gls{interactive} frame rates (we say that a method is interactive when the computation time for planning is less than the
time for execution). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

### 2017-01-16 body/overview.tex
updates

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--$\mathcal{P}_2$).
To achieve this the root path is first decomposed into a sequence of discrete root configurations, according to a user-defined discretization step.
Each root configuration is then extended into a whole-body configuration in static equilibrium.
The algorithm thus proceeds iteratively, starting from the whole-body initial configuration of the robot.
It takes advantage of the fact that each root configuration is fixed to generate the contact by considering each limb individually. 
%~ Candidate limb configurations resulting in the end-effector being close to the environment are selected from a database computed offline independently from the environment).
%~ Then we select one of these configurations based on user-defined heuristics (Figure~\ref{fig:contact_gen}), detailed in Appendix~\ref{sec:heuristics}.

@@

\subsection{Generating a discrete sequence of contact configurations --- \Pb}

>>>>

\subsection{Generating a discrete sequence of contact configurations --- \Pb}
The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--$\mathcal{P}_2$).
To achieve this the root path is first decomposed into a sequence of discrete root configurations, according to a user-defined discretization step.
Each root configuration is then extended into a whole-body configuration in static equilibrium.
The algorithm thus proceeds iteratively, starting from the whole-body initial configuration of the robot.
It takes advantage of the fact that each root configuration is fixed to generate the contact by considering each limb individually. 
%~ Candidate limb configurations resulting in the end-effector being close to the environment are selected from a database computed offline independently from the environment).
%~ Then we select one of these configurations based on user-defined heuristics (Figure~\ref{fig:contact_gen}), detailed in Appendix~\ref{sec:heuristics}.

@@

### 2017-01-16 body/rbprm.tex
updates

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces in contact. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

>>>>

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

### 2017-01-16 body/results.tex
updates

As it can be expected, the scaling results in a high decrease of the sensitivity, with an decrease of the specificity.
For HRP-2 we decided to set $s^*=1.2$.
%~ , and was also effective for the car egress scenario.

@@

For HyQ, there is a more uniform repartition of the times for the rubble scenario. For the obstacle scenario,
we observe a large increase of the time spent performing collision checking and inverse kinematics. This is explained
by the complexity of crossing the bridge: the robot has to stand high and create contacts close to each other because
the bridge is narrow. In many cases this results in collisions, thus a large number of contacts have to be evaluated.

>>>>

As it can be expected, the scaling results in a high decrease of the sensitivity, with a decrease of the specificity.
For HRP-2 we decided to set $s^*=1.2$.
%~ , and was also effective for the car egress scenario.

@@

On the other hand for HyQ most of the time is spent testing the static equilibrium of the candidates configurations..
